\section*{A}
%\rule{\textwidth}{0.2pt} % Thin horizontal line
\noindent \textbf{AI (Artificial Intelligence)}\\
\noindent A branch of computer science focused on creating systems capable of performing tasks that typically require human intelligence.

\vspace{4mm}

\noindent \textbf{API}\\
\noindent \textit{Application Programming Interface}: A set of rules and protocols for two or more software systems to communicate with each other, enabling the integration of their features and data. A documentation or standard that describes how such a connection or interface can be established or used is called an API specification.

\vspace{4mm}

\noindent \textbf{Autonomous System}\\
\noindent A network of computer-controlled operations or processes that perform tasks and make decisions without human intervention, using algorithms and often incorporating elements of artificial intelligence and machine learning.  generative art, it refers to a computer program or algorithm that independently creates art based on predefined rules or parameters, with minimal human intervention.

\section*{C}
\noindent \textbf{Climate Change}\\
\noindent Long-term changes in temperature and weather patterns, mainly caused by human activities like fossil fuel combustion, deforestation, and agriculture.

\section*{D}
\noindent \textbf{Data Bias}\\
\noindent The tendency of AI models to reflect biases present in their training data, leading to skewed or stereotypical outcomes.

\vspace{4mm}

\noindent \textbf{Generative Art}\\
\noindent The practice of using autonomous systems, often including computers or digital technology, to create art or become the artwork itself.

\section*{E}
\noindent \textbf{Extreme Weather Events}\\
\noindent Rare meteorological phenomena that significantly deviate from the historical averages of a particular place and time.

\section*{F}
\noindent \textbf{Framing}\\
\noindent The process of selecting and emphasising certain aspects of a topic while excluding others, thereby shaping the audience's perception.

\section*{I}
\noindent \textbf{Iconic Imagery}\\
\noindent Iconic imagery describes powerful, easily recognisable images that have become symbols or representations of broader concepts, themes, or issues. These images are often so strongly associated with certain subjects that they can instantly evoke specific ideas, emotions, or contexts in the viewer's mind.


\section*{M}
\noindent \textbf{Machine Learning}\\
\noindent A subset of AI that involves training computers to learn from and make decisions based on data.

\section*{N}
\noindent \textbf{News Factor Theory/ News Value Theory}\\
\noindent A theory in communication and media studies that identifies and analyses the elements making a story newsworthy. It suggests that certain characteristics, such as conflict, prominence, timeliness, proximity, and human interest, increase the likelihood of an event being reported as news. The theory was first introduced by \textcite{Galtung1965}.

\vspace{4mm}

\noindent \textbf{NGO}\\
\noindent \textit{Non-Governmental Organisation}: An independent organisation that operates outside of government control, often focused on addressing social, environmental, political, or human right issues. NGOs can range from small, local groups to large, international organisations.

\section*{P}
\noindent \textbf{Prompt Engineering}\\
\noindent The technique of crafting input prompts to guide AI models to produce desired outputs.

\section*{R}
\noindent \textbf{Responsible AI}\\
\noindent AI practices that prioritise ethical considerations, transparency, accountability, and fairness in AI systems and algorithms.

\section*{S}
\noindent \textbf{Saturation}\\
\noindent Saturation refers to the intensity or purity of a colour. A highly saturated colour is vivid and rich, while a less saturated colour appears more muted and greyish.

\section*{T}
\noindent \textbf{Text-to-Image Models}\\
\noindent AI technologies that convert textual descriptions into visual imagery, like \textit{OpenAI}'s \textit{DALL-E} and \textit{Midjourney}.

\vspace{4mm}

\noindent \textbf{Transformer-Based Architecture}\\
\noindent A neural network design used primarily in natural language processing, known for its self-attention mechanism.

\section*{V}
\noindent \textbf{Visual Attributes/Visual Modalities}\\
\noindent The characteristics of visual elements in an image, such as colour, brightness, and saturation, which can influence perception and emotional response.

\vspace{4mm}

\noindent \textbf{Visual Communication}\\
\noindent The conveyance of ideas and information through visual elements such as images, signs, and typography.

\vspace{4mm}

\noindent \textbf{Visual Synecdoche}\\
\noindent A visual representation that uses part of a whole to symbolise a larger concept or idea, often used in climate change communication.
